Structured Versus Decorated Cospans

### 2020-11-19 cospans.tex
fixed mistake in comment
missing citations on monoidal bicategories stuff
further small edits

\begin{thm}\label{thm:SC}
Given categories $\X$ and $\A$ with finite colimits and $L \maps \A \to \X$ a functor preserving
finite colimits, there is a symmetric monoidal double category $_L\lCsp(\X)$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical morphism is a morphism of $\A$,
\item a horizontal 1-cell from $a$ to $b$ is a cospan in $\X$ of the form
\begin{displaymath}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B);
\end{tikzpicture}
\end{displaymath}
\item a 2-morphism is a map of cospans in $\X$ of the form
\begin{displaymath}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\node (A') at (0,-1) {$L(a')$};
\node (B') at (1,-1) {$x'$};
\node (C') at (2,-1) {$L(b')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B)
(A') edge node[above]{$$} (B')
(C') edge node[above]{$$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [right]{$L(g)$} (C');
\end{tikzpicture}
\end{displaymath}
\end{itemize}
The symmetric monoidal structure is formed by using binary coproducts and the initial object.
\end{thm}

\begin{proof}
This was proved in \cite[Theorems~2.3 \& 3.9]{BC2}. Composition is formed by taking pushouts in $\X$ and identity horizontal 1-cells are $L(a)\xrightarrow{\raisebox{-3pt}[0pt][0pt]{=}}L(a)\xleftarrow{\raisebox{-3pt}[0pt][0pt]{=}}L(a)$. In particular, the double category structure only requires that $\X$ have pushouts, whereas the symmetric monoidal structure also requires that $\X$ has binary coproducts and an initial object (and thus all finite colimits), and that $\A$ has and $L$ preserves finite coproducts \cite[Theorem~3.2.3]{CourserThesis}.
\end{proof}

The following theorem establishes the isomorphism between structured and decorated cospan double categories under certain conditions.  We denote by $\Rex$ the 2-category of finitely cocomplete categories, finitely cocontinuous functors and natural transformations.   (In \cite{BC} we took as objects finitely cocomplete categories with chosen finite colimits, just so we could turn these into symmetric monoidal categories without any further choices!!!  Should we do that here???)

@@

These 2-equivalences further restrict to the case when the Grothendieck category $(X,\otimes)$ is specifically cocartesian monoidal itself. In that case, opfibrations
$(\X,+)\to(\A,+)$ that strictly preserve chosen coproducts and initial object
%, and such that $+$ preserves cocartesian liftings, Christina: this should ALWAYS be the case, by universal properties!
bijectively correspond to pseudofunctors into the 2-category of cocartesian categories; for more details, see \cite[Corollary 4.7]{MV} and the related discussion.

Finally, we can further restrict to the setting of opfibrations that also preserve pushouts, i.e.\ all finite colimits, which is important for our purposes. The following result is in fact more general since it relates the existence of any class of colimits in the total category of an opfibration to their existence in the fibres; for a more detailed exposition, see \cite[Corollary~4.9]{Hermida1999}.
%as well as Remark~4.11 therein regarding the non-necessary strictness condition. Is this obscuring or clarifying?

@@

Notice a subtlety in the above formulation: \cref{lem:MonGroth} provides a specifically induced monoidal structure on the fibres which becomes \cref{eq:explicitstructure1} in the cocartesian base case. If that fibrewise monoidal structure happens to itself be cocartesian, then it is necessarily isomorphic to \emph{any} other cocartesian structure on the fibres, by universal property of colimits. Therefore, we can simplify the above assumption to the one of the very \cref{thm:equiv} without having to specify that the structures came from monoidality of the pseudofunctor. However, we \emph{will} use the formulation of coproducts and initial object in terms of the monoidal structure of $F$, and that is why correspondence \cref{monGroth} is necessary.
DOES THIS MAKE SENSE TO ALL?? %still, we ARE going to use the formation of coproducts as above. Otherwise someone might think that the monoidal Groth story was redundant here!

@@

IN THE PROOF ABOVE, WE ACTUALLY POTENTIALLY NEED SYMMETRY. DOES IT FOLLOW?

@@

%\begin{itemize}
%\item 
%for every pair of objects $a$ and $b$ of $\mathbf{A}$, we have natural isomorphisms
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$\mathbf{A}(a,b)$};
%\node (B) at (2,0) {$\mathbf{X}(F(a),F(b))$};
%\node (C) at (0,-1) {$\mathbf{X}(G(a),G(b))$};
%\node (D) at (2,-1) {$\mathbf{X}(F(a),G(b))$};
%\node (E) at (1,-0.5) {$\sigma_{a,b} \Nearrow$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$F$} (B)
%(B) edge node[right]{$(\sigma_{b})_*$} (D)
%(A) edge node[left]{$G$} (C)
%(C) edge node[above]{$(\sigma_a)^*$} (D);
%\end{tikzpicture}
%\]
%where $(\sigma_a)^*$ and $(\sigma_b)_*$ are the functors induced by precomposition and postcompositon, respectively. Thus 
%for each morphism $f \maps a \to b$ in $\mathbf{A}$, an invertible natural 2-morphism $\sigma_f \maps G(f) \sigma_a \xrightarrow{\sim} \sigma_{b} F(f)$ in $\mathbf{B}$
%:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$F(a)$};
%\node (B) at (1,0) {$F(b)$};
%\node (C) at (0,-1) {$G(a)$};
%\node (D) at (1,-1) {$G(b)$};
%\node (E) at (0.5,-0.5) {$\sigma_f \Nearrow$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$F(f)$} (B)
%(B) edge node[right]{$\sigma_{b}$} (D)
%(A) edge node[left]{$\sigma_a$} (C)
%(C) edge node[above]{$G(f)$} (D);
%\end{tikzpicture}
%\]
%compatible with composition and identities.
%end{itemize}
%such that for each composable pair of morphisms $f \maps a \to b$ and $g \maps b \to c$ of $\mathbf{A}$, the following diagrams commute:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0.5) {$(G(g)G(f))\sigma_a$};
%\node (A') at (2,0.5) {$G(g)(G(f) \sigma_a)$};
%\node (B) at (0,-0.5) {$G(gf) \sigma_a$};
%\node (C) at (4,0.5) {$G(g) (\sigma_b F(f))$};
%\node (C') at (4,-0.5) {$\sigma_c F(gf)$};
%\node (D) at (6,0.5) {$(G(g) \sigma_b) F(f)$};
%\node (D') at (8,-0.5) {$\sigma_c (F(g)F(f))$};
%\node (F) at (8,0.5) {$(\sigma_c F(g)) F(f)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$a'$} (A')
%(A) edge node[left]{$\psi 1_{\sigma_a}$} (B)
%(A') edge node[above]{$1_{G(g)} \sigma_f$} (C)
%(B) edge node[above]{$\sigma_{gf}$} (C')
%(C) edge node [above] {${a'}^{-1}$} (D)
%(D') edge node [above] {$1_{\sigma_c} \phi$} (C')
%(D) edge node [above] {$\sigma_g 1_{F(f)}$} (F)
%(F) edge node [right] {$a'$} (D');
%\end{tikzpicture}
%\]
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$1_{G(a)} \sigma_a$};
%\node (B) at (1,1) {$G(1_a) \sigma_a$};
%\node (C) at (2,0) {$\sigma_a F(1_a)$};
%\node (D) at (0.5,-1) {$\sigma_a$};
%\node (E) at (1.5,-1) {$\sigma_a 1_{F(a)}$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[left]{$\psi 1_{\sigma_a}$} (B)
%(B) edge node[right]{$\sigma_{1_a}$} (C)
%(A) edge node [left] {$\ell'$} (D)
%(D) edge node [above] {${\rho'}^{-1}$} (E)
%(E) edge node [right] {$1_{\sigma_a} \phi$} (C);
%\end{tikzpicture}
%\]
%Christina: I removed the above, they are not needed I believe. Check again in the end.

>>>>

\begin{thm}\label{thm:SC}
Given categories $\X$ and $\A$ with finite colimits and $L \maps \A \to \X$ a functor preserving
finite colimits, there is a symmetric monoidal double category $_L\lCsp(\X)$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical morphism is a morphism of $\A$,
\item a horizontal 1-cell from $a$ to $b$ is a cospan in $\X$ of the form
\begin{displaymath}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\end{displaymath}
\item a 2-morphism is a map of cospans in $\X$ of the form
\begin{displaymath}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\node (A') at (0,-1) {$L(a')$};
\node (B') at (1,-1) {$x'$};
\node (C') at (2,-1) {$L(b')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B)
(A') edge node[above]{$$} (B')
(C') edge node[above]{$$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [right]{$L(g)$} (C');
\end{tikzpicture}
\end{displaymath}
\end{itemize}
Composition is formed by taking pushouts in $\X$ and identity horizontal 1-cells are $L(a)\xrightarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}L(a)\xleftarrow{\raisebox{-3pt}[0pt][0pt]{$=$}}L(a)$. 
Moreover, the vertical category is cocartesian monoidal $(\A,+,0_A)$ and the tensor of horizontal 1-cells is
\begin{displaymath}
\begin{tikzcd}[column sep={.4in,between origins}, row sep=.1in]
& x &&&& x' &&&& x+x' & \\
&&&\ot&&&& = &&&& \\
L(a)\ar[uur,"i"]&& L(b)\ar[uul,"o"'] && L(a')\ar[uur,"i'"] && L(b')\ar[uul,"o'"'] && L(a+a')\ar[uur,"i+i'"] && L(b+b')\ar[uul,"o+o'"'] 
\end{tikzcd}
\end{displaymath}
using that $L$ preserves coproducts, with monoidal unit $L(0_A)\xrightarrow{\raisebox{-3pt}[0pt][0pt]{$\sim$}}0_X\xleftarrow{\raisebox{-3pt}[0pt][0pt]{$\sim$}}L(0_A)$. %!!! {\chris Shouldn't the horizontal monoidal unit be the identity on the vertical monoidal unit?}
\end{thm}

\begin{proof}
This was proved in \cite[Theorems~2.3 \& 3.9]{BC2}. In particular, the double category structure only requires that $\X$ have pushouts, whereas the symmetric monoidal structure also requires that $\X$ has binary coproducts and an initial object (and thus all finite colimits), and that $\A$ has and $L$ preserves finite coproducts \cite[Theorem~3.2.3]{CourserThesis}. 
\end{proof}

The following theorem establishes the isomorphism between structured and decorated cospan double categories under certain conditions.  We denote by $\Rex$ the 2-category of finitely cocomplete categories, finitely cocontinuous functors and natural transformations.   (In \cite{BC} we took as objects finitely cocomplete categories with chosen finite colimits, just so we could turn these into symmetric monoidal categories without any further choices!!!  Should we do that here???) {\chris Can you tell if we can get away without doing that?}

@@

These 2-equivalences further restrict to the case when the Grothendieck category $(X,\otimes)$ is specifically cocartesian monoidal itself. In that case, opfibrations $(\X,+)\to(\A,+)$ that (strictly) preserve (chosen) coproducts and initial object
%, and such that $+$ preserves cocartesian liftings, Christina: this should ALWAYS be the case, by universal properties!
bijectively correspond to pseudofunctors into the 2-category of cocartesian categories; for more details, see \cite[Corollary 4.7]{MV} and the related discussion.

Finally, we can further restrict to the setting of opfibrations that also preserve pushouts, i.e.\ all finite colimits, which is important for our purposes. The following result is in fact more general since it relates the existence of any class of colimits in the total category of an opfibration to their existence in the fibres; for a more detailed exposition, as well as the non-necessary strictness condition, see \cite[Corollary~4.9\&Remark~4.11]{Hermida1999}. %Is this obscuring or clarifying?

@@

Notice a subtlety in the above formulation: \cref{lem:MonGroth} provides a specifically induced monoidal structure on the fibres which becomes \cref{eq:explicitstructure1} in the cocartesian base case. If that fibrewise monoidal structure happens to itself be cocartesian, then it is necessarily isomorphic to \emph{any} other cocartesian structure on the fibres, by universal property of colimits. Therefore, we can simplify the above assumption to the one of the very \cref{thm:equiv} without having to specify that the structures came from monoidality of the pseudofunctor. However, we \emph{will} use the formulation of coproducts and initial object in terms of the monoidal structure of $F$, and that is why correspondence \cref{monGroth} is necessary.
DOES THIS MAKE SENSE TO ALL??

@@

{\chris In the proof above, we probably need isomorphism. Does it follow?}

@@

{\chris There used to be references for more details on the material that follows. Why would we not provide them?}
